Accurate binding of calcium to phospholipid bilayers by effective inclusion of electronic polarization

### 2017-10-06 Manuscript/manuscript.tex
Pavel's comments and corrections fully implemented into manuscript. This commit: Conclusions.

\section{Conclusions}
We show that the \ce{Na^+} and \ce{Ca^2+} binding in phospholipid bilayers can
be accurately described with classical MD simulation models, where electronic
polarization is effectively included by using electronic continuum correction (ECC)~\cite{leontyev11}.
This is a significant improvement over other available lipid models,
which all overestimate specific cation binding affinities~\cite{catte16}.  
The newly proposed model, which we denote as ''ECC-lipids~17'', 
exhibits accurate head group order parameter response to
bound cations, monovalent \ce{Na^+} and cationic surfactant dihexadecyldimethylammonium bromide, 
and divalent \ce{Ca^2+}
also quantifying their binding affinities.
Moreover, ECC-lipids~17 reproduce the lipid bilayer structural details
with similar accuracy as other state of the art lipid models~\cite{catte16}.
Several water models 
(OPC3\cite{Izadi16}, OPC \cite{Izadi14}, SPC/E \cite{Berendsen1987} and TIP4p/2005 \cite{Abascal2005}) 
were used to exemplify the transferability of 
the parameters of the new ECC-lipids~17 force field.

Direct analysis of calcium binding details from MD simulations is in agreement
with ternary complex model, which is suggested based on NMR data \cite{altenbach84}.
In this model 1~calcium binds to 2~POPC molecules, which together form a ternary
complex.
\todo{Continue summary using previous section once it is finished.}

The electronic continuum correction is applied here on Lipid14 POPC model \cite{dickson14},
but we expect that the correction can be generalized also for other lipids
and force fields.
The parameters can be used with existing standard nucleic acid and protein force fields, e.g. AMBER-FB15~\cite{Wang2017}. 
We suggest using state of the art water models like OPC3\cite{Izadi16} or OPC \cite{Izadi14},
which yield higher accuracy than the traditional TIP3p water model~\cite{jorgensen83}.

@@

% If you have acknowledgments, this puts in the proper section head.
\begin{acknowledgments}
% Put your acknowledgments here.
\end{acknowledgments}
\newpage
\appendix
\begin{center}
{\bf SUPPLEMENTARY INFORMATION}
\end{center}

@@

\begin{figure}[]
  \centering
  \includegraphics[height=9.0cm,angle=-90]{../Fig/bound-CAs_conc-eccl17.eps}
  \caption{\label{fig:cacl-bind}
    Ternary complex binding model of \ce{Ca^{2+}} to a POPC membrane 
    that assumes the stoichiometry of 2~POPC:1~\ce{Ca^{2+}} (details in reference~\citenum{altenbach84}) 
    provides a good fit to experimental measurements~\cite{altenbach84}
    and it also provides a good fit to our simulation data. 
    Note that the units in the reference~\citenum{altenbach84} are different from the units presented here,
    and, hence, the observed slope of the linear relationship is different.
    }
\end{figure}

\begin{figure*}[tbp]
  \centering
  \includegraphics[width=8.0cm]{../Fig/ipython_nb/PN_angle_OrdPars-A-B-g3_L14-ECCL17_q80_sig89_CaCl_waterModels.pdf}
  \caption{\label{fig:ordPars_waterModels}
    Changes of head group order parameters of POPC bilayer as a function of CaCl$_2$ concentrations
    are shown from simulations with different force fields and water models together with experimental data 
    (DPPC \cite{akutsu81} and POPC \cite{altenbach84}). 
    Ion concentrations in bulk water are shown in x-axis. 
    Values from simulations are calculated from the of cation number density $C_{np}$
    from the region at the simulatin box edge with the constant ion concentration as [ion]=$C_{np}/0.602$.
  }
\end{figure*}

>>>>

\section{Conclusions}
We show that the \ce{Na^+} and \ce{Ca^2+} binding to phospholipid bilayers can
be accurately described with classical force field MD simulation models, where electronic
polarization is effectively included using the electronic continuum correction (ECC)~\cite{leontyev11}.
This is a significant improvement over other available lipid models,
which all overestimate specific cation binding affinities~\cite{catte16}.  
The newly proposed model, which we denote as ''ECC-lipids'', 
exhibits accurate head group order parameter response to
bound cations, namely monovalent \ce{Na^+} and cationic surfactant dihexadecyldimethylammonium bromide, 
and divalent \ce{Ca^2+},
allowing also for quantifying their binding affinities.
At the same time, ECC-lipids reproduce the lipid bilayer structural details
with similar accuracy as other state of the art lipid models~\cite{catte16}.
Several water models 
(OPC3\cite{Izadi16}, OPC \cite{Izadi14}, SPC/E \cite{Berendsen1987} and TIP4p/2005 \cite{Abascal2005}) 
were used to exemplify the transferability of 
the parameters of the new ECC-lipids force field.

Direct analysis of the details of calcium binding from MD simulations is in agreement
with ternary complex model, which is suggested based on NMR data \cite{altenbach84}.
In this model one calcium binds to two POPC molecules, which together form a ternary
complex.
\todo{Continue summary using previous section once it is finished.}

The electronic continuum correction is built here upon the Lipid14 POPC model \cite{dickson14},
but we expect that the correction can be used also for other lipids
and force fields.
\todo{JOE: Do we want to suggest that ECC-lipids can be used 
with current force fields to study for example membrane proteins? 
This was expressed with the following sentence: 
The parameters can be used with existing standard nucleic acid and protein force fields, e.g. AMBER-FB15~\cite{Wang2017}. }
We suggest using state of the art water models like OPC3\cite{Izadi16} or OPC \cite{Izadi14},
which yield higher accuracy than the traditional TIP3p water model~\cite{jorgensen83}.

@@

% If you have acknowledgments, this puts in the proper section head.
\begin{acknowledgments}
% Put your acknowledgments here.
P.J. acknowledges support from the Czech Science Foundation (grant no. 16-01074S) 
and 600 from the Academy of Finland via the FiDiPro award.
\end{acknowledgments}

\newpage
\appendix

\begin{center}
{\bf SUPPLEMENTARY INFORMATION}
\end{center}

@@

\begin{figure}[tbp]
  \centering
  \includegraphics[height=9.0cm,angle=-90]{../Fig/bound-CAs_conc-eccl17.eps}
  \caption{\label{fig:cacl-bind}
    Ternary complex binding model of \ce{Ca^{2+}} to a POPC membrane 
    that assumes the stoichiometry of 2~POPC:1~\ce{Ca^{2+}} (details in reference~\citenum{altenbach84}) 
    provides a good fit to experimental measurements~\cite{altenbach84}
    and it also provides a good fit to our simulation data. 
    Note that the units in the reference~\citenum{altenbach84} are different from the units presented here,
    and, hence, the observed slope of the linear relationship is different.
    }
\end{figure}

\begin{figure}[tbp]
  \centering
  \includegraphics[width=8.0cm]{../Fig/ipython_nb/PN_angle_OrdPars-A-B-g3_L14-ECCL17_q80_sig89_CaCl_waterModels.pdf}
  \caption{\label{fig:ordPars_waterModels}
    Changes of head group order parameters of POPC bilayer as a function of CaCl$_2$ concentrations
    are shown from simulations with different force fields and water models together with experimental data 
    (DPPC \cite{akutsu81} and POPC \cite{altenbach84}). 
    Ion concentrations in bulk water are shown in x-axis. 
    Values from simulations are calculated from the of cation number density $C_{np}$
    from the region at the simulatin box edge with the constant ion concentration as [ion]=$C_{np}/0.602$.
  }
\end{figure}
